Supplementary Information 1

### 2025-05-21 index.Rmd
last changes in html

---
title: "Supplementary Information 1"
subtitle: "Assessing ecosystem effects on the stock productivity of Antarctic Krill (*Euphausia superba*): An Integrated modeling perspective"
author: "Mardones, M; Jarvis Mason, E.T.; Pinones, A.;  Santa Cruz, F.; Cárdenas, C.A"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: SA_krill.bib
#csl: apa.csl
csl: icesjournal.csl
link-citations: yes
linkcolor: blue
output:
  bookdown::html_document2:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: cosmo
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
    number_sections: false
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup1, echo=FALSE}
#rm(list = ls())
#set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 6,
                      fig.width = 7,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      fig.path = "Figs/",
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

```{r message=FALSE, eval=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) # facilita la ejecución paralela en R
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
library(tidyr)
library(mixR)
library(readxl)
library(tidyverse)
library(ggridges)
library(ggrepel)
library(broom)
library(lmtest)
library(flextable)
```

```{r message=FALSE, warning=FALSE, echo=TRUE}
# dir01 <- here("s01") # agreggate data (no spatial diferences)
# dir1<-here("s1") # Data strata flishery
# dir2<-here("s2") # Same 9 with areas (SubStrata) as fleet. Dif size comoposition and dif CPUE and dif survey length and biomass data by strata
# dir3<-here("s3") # without S-R
# dir4<-here("s4") # 
# dir5<-here("s5") # 
# dir6<-here("s6") # 
# dir7<-here("s7") # 2 set parametres EMM-2024/23 (Mardones)
# dir8<-here("s8") # s1 platoons
# dir9<-here("s9") # s1 w/ blocks
dir1.1<-here("s1.1") # sin predador ni ambiental
dir1.2 <- here("s1.2") # s1.1 C/ predador
dir1.3 <- here("s1.3") # s1.1 solo env
dir1.4 <- here("s1.4") # s1.2 predator and env
Figs <- here("Figs")# S
```

@@

This study aims to evaluate the impact of ecosystem components—such as environmental variables and predator-prey interactions—on the productivity and key population dynamics of *Euphausia superba* in Subarea 48.1. By incorporating these factors into the assessment, we analyze how krill population variables respond to ecological variability, providing insights into their resilience and potential management implications.

Here, the **reference model** represents a baseline assessment of *Euphausia superba* population dynamics in Subarea 48.1, excluding environmental and ecological variables. This model assumes that krill productivity and population parameters are driven  by intrinsic biological processes, such as growth, mortality, and recruitment and fishery impacts without accounting for external influences like environmental variability or predation pressure. By serving as a *control scenario*, this model provides a benchmark against which the impact of ecosystem components in productivity can be evaluated, allowing for a direct comparison of how environmental and ecological factors influence krill stock dynamics.

The study area corresponds to Subarea 48.1 in the WAP, where most krill fishing activity occurs and where CCAMLR has been working to implement a new management strategy at a reduced spatial scale (Figure \@ref(fig:mapa)). To refine the spatial resolution of krill population dynamics and analysis, we adopted the five strata proposed by CCAMLR: Bransfield Strait, Elephant Island, Gerlache Strait, Joinville Island, and the Southwest [@Dornam2021; REF] (Fig 1 S1). While these strata were originally designed for management purposes without considering specific biological or population characteristics, their subdivision enables us to analyze the data at a finer scale. This level of detail facilitates the identification of heterogeneity across various data sources, whether these originate from the fishery or from environmental survey data, providing a more nuanced understanding of krill population dynamics. This spatial segregation of information is incorporated into the stock assessment model, making it a spatially explicit model that accounts for regional variability in population dynamics and external factors.

@@

In this approach, spatial structure is incorporated implicitly by treating different areas as separate fleets [@Nielsen2021; @Waterhouse2014]. This constitutes a spatially implicit modeling framework, where differences among strata are recognized both from the perspective of krill population dynamics and from the influence of environmental variability within Subarea 48.1 (Figure \@ref(fig:concem)).

```{r concem, out.width='40%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
knitr::include_graphics('Figs/conceptual.jpeg')
```

@@

Stock Synthesis (v.3.30.21)  is a widely used tool for assessing fish and invertebrate populations, including Antarctic krill. SS3 is implemented in `C++` with estimation enabled through automatic differentiation (ADMB) [@Fournier2012; @Methot2013]. In this exercise, SS3 is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers. The model simulates population processes such as growth, maturity, fecundity, recruitment, movement, and mortality, while also integrating environmental variability and predator-prey relationships to refine estimates of population trends. The analysis of model outputs is conducted using R, utilizing the *r4ss* and *ss3diags* packages [@Taylor2019; @Winker2023]. By leveraging a spatially implicit, ecosystem-informed approach, this assessment provides a robust framework for evaluating krill stock dynamics under changing environmental conditions. These insights are crucial for informing sustainable management strategies in the Antarctic Peninsula region, where krill plays a foundational role in the marine food web.

### Age Key Length (ALK)

In a catch-at-length model like krill assessment the AKL matrix is modelled trought parametrization process

@@

## Parameters

@@

```{r}
parInit %>%
  mutate(across(where(is.numeric), ~ round(., 3)))  %>% 
  kbl(booktabs = TRUE,
      format = "html",
      position = "ht!",
      caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
  kable_paper("hover", full_width = TRUE) %>%  # Expande el ancho
  kable_styling(latex_options = c("striped", "condensed"),
                full_width = TRUE,  # Expande la tabla al ancho completo
                font_size = 12,
                html_font = "Times New Roman") %>%  # Aplica la fuente Times New Roman
  pack_rows(index = c("Natural Mortality" = 1,
                      "Growth" = 5,
                      "relationship Length-Weigth" = 2,
                      "Maturity" = 2,
                      "S-R relation" = 5,
                      "Catchability" = 2,
                      "Selectivity" = 4))

```
Source of data input.

Ibdex abundance in Figure \@ref(fig:index)

```{r}
dat1_std <- dat1$CPUE %>%
  mutate(
    index_std = (obs - min(obs)) / (max(obs) - min(obs)),
    index = recode(index,
                   `1` = "FISHERYBS",
                   `2` = "FISHERYEI",
                   `3` = "FISHERYGS",
                   `4` = "FISHERYJOIN",
                   `5` = "FISHERYSSIW",
                   `6` = "SURVEYBS",
                   `7` = "SURVEYEI",
                   `8` = "SURVEYGS",
                   `9` = "SURVEYJOIN",
                   `10` = "SURVEYSSIW",
                   `11` = "PREDATOR"),
    index = factor(index, levels = c("FISHERYBS", 
                                     "FISHERYEI", 
                                     "FISHERYGS",
                                     "FISHERYJOIN", 
                                     "FISHERYSSIW", 
                                     "SURVEYBS", 
                                     "SURVEYEI", 
                                     "SURVEYGS", 
                                     "SURVEYJOIN", 
                                     "SURVEYSSIW", 
                                     "PREDATOR")),
    category = case_when(
      index %in% c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW") ~ "Fishery",
      index %in% c("SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "SURVEYSSIW") ~ "Survey",
      index == "PREDATOR" ~ "Predator",
      TRUE ~ NA_character_  
    )
  )
```

```{r index, fig.width=9, fig.height=4, fig.cap="Standardized indices of krill index abundance and consumption from fishery-dependent, fishery-independent, and predator-based data sources across different strata within Subarea 48.1. Each panel represents a distinct spatial or functional stratum, with trend lines indicating temporal variation from 1990 to 2020. Colors denote data source categories: green for fishery, orange for scientific surveys, and purple for predator-based indices. These patterns highlight spatial and temporal heterogeneity in krill dynamics across the subarea."}

@@

```{r warning=FALSE}
dat_long <- dat1$lencomp %>% 
  pivot_longer(cols = starts_with("l"),  
               names_to = "Talla",  
               values_to = "Frecuencia") %>% 
  mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
  mutate(FltSvy = recode(FltSvy,
                         `1` = "FISHERYBS",
                         `2` = "FISHERYEI",
                         `3` = "FISHERYGS",
                         `4` = "FISHERYJOIN",
                         `5` = "FISHERYSSIW",
                         `6` = "SURVEYBS",
                         `7` = "SURVEYEI",
                         `8` = "SURVEYGS",
                         `9` = "SURVEYJOIN",
                         `10` = "SURVEYSSIW",
                         `11` = "PREDATOR")) %>%
  mutate(FltSvy = factor(FltSvy, levels = c("FISHERYBS", 
                                            "FISHERYEI", 
                                            "FISHERYGS",
                                            "FISHERYJOIN", 
                                            "FISHERYSSIW",
                                            "SURVEYBS",
                                            "SURVEYEI", 
                                            "SURVEYGS", 
                                            "SURVEYJOIN",
                                            "SURVEYSSIW",
                                            "PREDATOR")))

@@

@@

In Table 1 we have ten scenarios to test different option in modeling
about main consideration in assessment of krill population.

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|    s1.1  | Spatial data without environmental and predator components                           |
|    s1.2  | "s1.1" with predator components                                                      |
|    s1.3  | "s1.1" with environmental variable                                                  |
|    s1.4  | "s1.1" w/ both, predator fleet and environmental variable                           |

### Run Models

```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
directorios <- c("s1.1",
                 "s1.2",
                 "s1.3",
                 "s1.4")

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dir1.1,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)

# s1.2
r4ss::run(
  dir = dir1.2,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)

# s1.3
r4ss::run(
  dir = dir1.3,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)

# s1.4
r4ss::run(
  dir = dir1.4,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)
```

Data used en both (spatial and No spatial models) `s1.1` and `s1.4` to compare

```{r fig.height=10, fig.width=6}
par(mfrow=c(2,1), mar=c(5,4,2,1))
SSplotData(base.model1.1, 
           subplots = 1,
           ptsize = 0.5)

SSplotData(base.model1.4, 
           subplots = 1,
           ptsize = 0.5)
```

@@

@@

Main Variables poulation in `s1.1` scenario

```{r fig.height=2, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1)) 
SSplotTimeseries(base.model1.1,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.1,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.1,
                 subplot = 7,
                 xlab = "")
```

Main Variables poulation in `s1.2` scenario

```{r fig.height=2, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.2,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.2,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.2,
                 subplot = 7,
                 xlab = "")
```

Main Variables poulation in `s1.3` scenario

```{r fig.height=2, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.3,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.3,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.3,
                 subplot = 7,
                 xlab = "")
```

Main Variables poulation in `s1.4` scenario

```{r fig.height=2, fig.width=7}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.4,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.4,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.4,
                 subplot = 7,
                 xlab = "")
```
Selectivity

```{r}
par(mfrow=c(2,2), mar=c(5,4,1,1))
SSplotSelex(base.model1.1,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.1", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.2,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.2", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.3,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.3", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.4,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.4", side=3, line=0.5, cex=1.2)
```
```{r}
par(mfrow=c(2,2), mar=c(5,4,1,1))
SSplotBiology(base.model1.1,
            subplots = 1)
SSplotBiology(base.model1.2,
            subplots = 1)
SSplotBiology(base.model1.3,
            subplots = 1)
SSplotBiology(base.model1.4,
            subplots = 1)
```

Natural Mortality

```{r eval=FALSE}
par(mfrow=c(4,2), mar=c(5,4,2,1))
SSplotBiology(base.model1.1,
              subplots = 21)
SSplotBiology(base.model1.2,
            subplots = 21)
mtext("s1.2", side=3, line=0.5, cex=1.2)
SSplotBiology(base.model1.3,
            subplots = 21)
mtext("s1.3", side=3, line=0.5, cex=1.2)
SSplotBiology(base.model1.4,
            subplots = 21)
mtext("s1.4", side=3, line=0.5, cex=1.2)
```

```{r}
par(mfrow=c(2,2), mar=c(5,4,2,1))

SSplotIndices(base.model1.1, subplots = 9)
mtext("s1.1", side=3, line=0.5, cex=1.2)

SSplotIndices(base.model1.2, subplots = 9)
mtext("s1.2", side=3, line=0.5, cex=1.2)

SSplotIndices(base.model1.3, subplots = 9)
mtext("s1.3", side=3, line=0.5, cex=1.2)

SSplotIndices(base.model1.4, subplots = 9)
mtext("s1.4", side=3, line=0.5, cex=1.2)

>>>>

---
title: "Supplementary Information 1"
subtitle: "Assessing ecosystem effects on the stock productivity of Antarctic Krill (*Euphausia superba*): An Integrated modeling perspective"
author: "Mardones, M; Jarvis Mason, E.T.; Piñones, A.; Krüger, L.,Santa-Cruz, F.; Watters, G., Cárdenas, C.A; Methot, R."
date:  "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: SA_krill.bib
csl: apa.csl
#csl: icesjournal.csl
link-citations: yes
linkcolor: blue
output:
  bookdown::html_document2:
    keep_md: true
    toc: true
    toc_deep: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: simplex
    fontsize: 0.9em
    linestretch: 1.7
    html-math-method: katex
    self-contained: true
    code-tools: true
    number_sections: false
editor_options: 
  markdown: 
    wrap: 72
---

## MODELING SETTING

```{r setup1, echo=FALSE}
#rm(list = ls())
#set.seed(999)
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      fig.path = "Figs/",
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
#XQuartz is a mess, put this in your onload to default to cairo instead
options(bitmapType = "cairo") 
# (https://github.com/tidyverse/ggplot2/issues/2655)
# Lo mapas se hacen mas rapido
```

### Libraries

Libraries necessary to made this analysis;

```{r message=FALSE, eval=TRUE, echo=TRUE}
# install.packages("devtools")
# devtools::install_github("r4ss/r4ss", ref="development")
# install.packages("caTools")
# library("caTools")
# install.packages("r4ss")
library(r4ss)
library(here)
#remotes::install_github("PIFSCstockassessments/ss3diags")
library(ss3diags)
library(kableExtra)
library(doParallel) 
detectCores()
registerDoParallel(8)
library(ggpubr)
library(tibble)
library(openxlsx)
library(ggthemes)
library(forecast)
library(tidyr)
library(mixR)
library(readxl)
library(tidyverse)
library(ggridges)
library(ggrepel)
library(broom)
library(lmtest)
library(flextable)
library(car)
```

The repository with files templates by scenario to replicate this analysis can be found in this [GitHub author link](https://github.com/MauroMardones/SA_Krill/tree/main/scenarios)

```{r message=FALSE, warning=FALSE, echo=TRUE}
dir1.1<-here("s1.1") # no predador- environmental
dir1.2 <- here("s1.2") # s1.1 w/ predador
dir1.3 <- here("s1.3") # s1.1 w/ env
dir1.4 <- here("s1.4") # s1.1 predator and env
Figs <- here("Figs")
```

```{r eval=FALSE, message=F, include=FALSE, echo=TRUE}
### Run Models
directorios <- c("s1.1",
                 "s1.2",
                 "s1.3",
                 "s1.4")

```{r eval=FALSE, message=F, include=FALSE}
#OR with rby separate
#shell(cmd="ss") # run SS to windows
# or 
r4ss::run(
  dir = dir1.1,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)
# s1.2
r4ss::run(
  dir = dir1.2,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)
# s1.3
r4ss::run(
  dir = dir1.3,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)
# s1.4
r4ss::run(
  dir = dir1.4,
  exe = "ss_osx",
  skipfinished = FALSE, 
  show_in_console = TRUE 
)
```

@@

This study aims to evaluate the impact of ecosystem components—such as environmental variables and predator-prey interactions—on the productivity and key population dynamics of *Euphausia superba* in Subarea 48.1. By incorporating these factors into the assessment, we analyze how krill population variables respond to ecological variability, providing insights into their resilience and potential management implications. This code forms part of the supplementary material for the scientific article titled **"Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective."**

## Spatial dimension of stock assessment

The study focuses on Subarea 48.1 in the Western Antarctic Peninsula, where most krill fishing occurs and CCAMLR is advancing spatially refined management. We use five CCAMLR-defined strata to increase spatial resolution, allowing detection of regional differences in krill dynamics (Figure \@ref(fig:mapa)). This structure, integrated into the stock assessment model, enhances its capacity to capture spatial variability in both biological and environmental data. In this approach, spatial structure is incorporated implicitly by treating different areas as separate fleets [@Nielsen2021; @Waterhouse2014]. This constitutes a spatially implicit modeling framework, where differences among strata are recognized both from the perspective of krill population dynamics and from the influence of environmental variability within Subarea 48.1.

@@

Figure \@ref(fig:conceptual) illustrates a conceptual model with the spatial distribution and movement of Antarctic krill in the Western Antarctic Peninsula (WAP). Adult krill are concentrated and move primarily in the northern areas, while juvenile krill dominate the southern regions. Arrows indicate directional flows, suggesting ontogenetic or environmentally-driven migration patterns. In terms of assumptions, although there are spatial differences in population structure, all these areas are considered part of a single, closed population unit upon which the stock assessment is conducted. By leveraging a spatially implicit, ecosystem-informed approach, this assessment provides a robust framework for evaluating krill stock dynamics under changing environmental conditions. These insights are crucial for informing sustainable management strategies in the Antarctic Peninsula region, where krill plays a foundational role in the marine food web.

```{r conceptual, out.width='50%', fig.cap="Conceptual model used to model dynamics population in Antarctic krill in WAP"}
knitr::include_graphics('Figs/conceptual.jpeg')
```

@@

Stock Synthesis (v.3.30.21)  is a widely used tool for assessing fish and invertebrate populations, including Antarctic krill. SS3 is implemented in `C++` with estimation enabled through automatic differentiation (ADMB) [@Fournier2012; @Methot2013]. The spurce code can be find in [Github SS3 repository](https://github.com/nmfs-ost/ss3-source-code). In this exercise, `SS3` is configured as an integrated stock assessment model, explicitly accounting for age and size structure while incorporating key ecosystem drivers. The model simulates population processes such as growth, maturity, fecundity, recruitment, movement, and mortality, while also integrating environmental variability and predator-prey relationships to refine estimates of population trends in krill. The analysis of model outputs is conducted using R, utilizing the *r4ss* and *ss3diags* packages [@Taylor2019; @Winker2023]. Integrated models can effectively capture the age structure by transforming length observations into population-level dynamics [@Lee2024; @Punt2013].

On the other hand, length data, which are cost-effective and readily available [@Canales2021; @Chong2019], provide valuable insights into krill population structure due to their correlation with age [@Thanassekos2014]. For made this transformation, the models often use an Age-Length Keys (ALKs). An ALK is a probabilistic matrix that estimates the likelihood of individuals of a given length belonging to specific age classed and this method assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases [@ICCA2003; @Punt2003; @Punt2013; @Lee2024]. The ALK matrix is typically defined as:

In a catch-at-length model like krill assessment the AKL matrix (Figure \@ref(fig:AKL)) is modeled trough parametrization process and have this shape;

@@

## Data

### Parameters

@@

```{r}
parInit %>%
  mutate(across(where(is.numeric), ~ round(., 3)))  %>% 
  kbl(booktabs = TRUE,
      format = "html",
      caption = "Input parameters for the initial SS3 model of krill. Each parameter line contains a minimum value (LO), maximum value (HI), and initial value (INIT). If the phase (PHASE) for the parameter is negative, the parameter is fixed as input") %>%
      kable_styling(latex_options = "scale_down",  # Expande la tabla al ancho completo
                font_size = 8) %>%  # Aplica la fuente Times New Roman
  pack_rows(index = c("Natural Mortality" = 1,
                      "Growth" = 5,
                      "relationship Length-Weigth" = 2,
                      "Maturity" = 2,
                      "S-R relation" = 5,
                      "Catchability" = 2,
                      "Selectivity" = 4))

### Index

Abundance index in Figure \@ref(fig:index)

```{r}
dat1_std <- dat1$CPUE %>%
  mutate(
    index_std = (obs - min(obs)) / (max(obs) - min(obs)),
    index = dplyr::recode(index,
                   `1` = "FISHERYBS",
                   `2` = "FISHERYEI",
                   `3` = "FISHERYGS",
                   `4` = "FISHERYJOIN",
                   `5` = "FISHERYSSIW",
                   `6` = "SURVEYBS",
                   `7` = "SURVEYEI",
                   `8` = "SURVEYGS",
                   `9` = "SURVEYJOIN",
                   `10` = "SURVEYSSIW",
                   `11` = "PREDATOR"),
    index = factor(index, levels = c("FISHERYBS", 
                                     "FISHERYEI", 
                                     "FISHERYGS",
                                     "FISHERYJOIN", 
                                     "FISHERYSSIW", 
                                     "SURVEYBS", 
                                     "SURVEYEI", 
                                     "SURVEYGS", 
                                     "SURVEYJOIN", 
                                     "SURVEYSSIW", 
                                     "PREDATOR")),
    category = case_when(
      index %in% c("FISHERYBS", "FISHERYEI", "FISHERYGS", "FISHERYJOIN", "FISHERYSSIW") ~ "Fishery",
      index %in% c("SURVEYBS", "SURVEYEI", "SURVEYGS", "SURVEYJOIN", "SURVEYSSIW") ~ "Survey",
      index == "PREDATOR" ~ "Predator",
      TRUE ~ NA_character_  
    )
  )
```

```{r index, fig.width=9, fig.height=6, fig.cap="Standardized indices of krill index abundance and consumption from fishery-dependent, fishery-independent, and predator-based data sources across different strata within Subarea 48.1. Each panel represents a distinct spatial or functional stratum, with trend lines indicating temporal variation from 1990 to 2020. Colors denote data source categories: green for fishery, orange for scientific surveys, and purple for predator-based indices. These patterns highlight spatial and temporal heterogeneity in krill dynamics across the subarea."}

@@

### Length comositions

```{r warning=FALSE}
dat_long <- dat1$lencomp %>% 
  pivot_longer(cols = starts_with("l"),  
               names_to = "Talla",  
               values_to = "Frecuencia") %>% 
  mutate(Talla = as.numeric(sub("l", "", Talla))) %>%  
  mutate(FltSvy = dplyr::recode(FltSvy,
                         `1` = "FISHERYBS",
                         `2` = "FISHERYEI",
                         `3` = "FISHERYGS",
                         `4` = "FISHERYJOIN",
                         `5` = "FISHERYSSIW",
                         `6` = "SURVEYBS",
                         `7` = "SURVEYEI",
                         `8` = "SURVEYGS",
                         `9` = "SURVEYJOIN",
                         `10` = "SURVEYSSIW",
                         `11` = "PREDATOR")) %>%
  mutate(FltSvy = factor(FltSvy, levels = c("FISHERYBS", 
                                            "FISHERYEI", 
                                            "FISHERYGS",
                                            "FISHERYJOIN", 
                                            "FISHERYSSIW",
                                            "SURVEYBS",
                                            "SURVEYEI", 
                                            "SURVEYGS", 
                                            "SURVEYJOIN",
                                            "SURVEYSSIW",
                                            "PREDATOR")))

@@

## Summary data sources to Modelling

Finally, this information and all sources can be represented through the following flow diagram (Figure \@ref(fig:path)) of inputs, model, and outputs.

Figure \@ref(fig:dataserie) show time series of differente componentes of data sources to this krill stock assessment.

```{r dataserie, fig.width=9, fig.height=7}
SSplotData(base.model1.4, 
           subplots = 1,
           ptsize = 0.5)
```

@@

Here, the **reference model** is `s1.1`, wich represents a baseline assessment of *Euphausia superba* population dynamics in Subarea 48.1, excluding environmental and ecological variables. This model assumes that krill productivity and population parameters are driven  by intrinsic biological processes, such as growth, mortality, and recruitment and fishery impacts without accounting for external influences like environmental variability or predation pressure. By serving as a *control scenario*, this model provides a benchmark against which the impact of ecosystem components in productivity can be evaluated, allowing for a direct comparison of how environmental and ecological factors influence krill stock dynamics.

Table \@ref(tab:scenarios) show scenarios used for modelling dynamics in krill

```{r scenario, eval=TRUE}
# Tu código de flextable aquí
scenario_table <- data.frame(
  Scenario = c("s1.1", "s1.2", "s1.3", "s1.4"),
  Description = c(
    "Spatial data without environmental and predator components",
    "\"s1.1\" with predator components",
    "\"s1.1\" with environmental variable",
    "\"s1.1\" with both predator fleet and environmental variable"
  ),
  stringsAsFactors = FALSE
)

scenario_table %>%
  kbl(caption = "Scenarios used for modelling dynamics in krill",
      align = "l",  # Left alignment (or use "c" for center)
      booktabs = TRUE,
      col.names = c("Scenario", "Description")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center",
                font_size = 10,  # Equivalent to fontsize(size = 10)
                latex_options = "HOLD_position") %>%
  row_spec(0, bold = TRUE, color = "black") %>%  # Header formatting
  column_spec(1, width = "2cm") %>%  # Adjust column widths
  column_spec(2, width = "10cm")
```

@@

## Population Variables by scenario

@@

Main Variables poulation in `s1.1` scenario (Figure \@ref(fig:scen1))

```{r scen1, fig.height=2, fig.width=7, fig.cap="Main variables scenario s1.1"}
par(mfrow=c(1,3), mar=c(5,4,1,1)) 
SSplotTimeseries(base.model1.1,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.1,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.1,
                 subplot = 7,
                 xlab = "")
```

Main Variables poulation in `s1.2` scenario (Figure \@ref(fig:scen2))

```{r scen2, fig.height=2, fig.width=7, fig.cap="Main variables scenario s1.2"}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.2,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.2,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.2,
                 subplot = 7,
                 xlab = "")
```

Main Variables poulation in `s1.3` scenario (Figure \@ref(fig:scen3))

```{r scen3, fig.height=2, fig.width=7, fig.cap="Main variables scenario s1.3"}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.3,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.3,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.3,
                 subplot = 7,
                 xlab = "")
```

Main Variables poulation in `s1.4` scenario (Figure \@ref(fig:scen4))

```{r scen4, fig.height=2, fig.width=7 , fig.cap="Main variables scenario s1.4"}
par(mfrow=c(1,3), mar=c(5,4,1,1))
SSplotTimeseries(base.model1.4,
                 subplot = 1,
                 xlab = "")
SSplotTimeseries(base.model1.4,
                 subplot = 9,
                 xlab = "")
SSplotTimeseries(base.model1.4,
                 subplot = 7,
                 xlab = "")
```

Selectivity estimated by scenario in Figure \@ref(fig:selectivity).

```{r selectivity, fig.height=8, fig.width=8, fig.cap="Selectivity by fleet in each scenario"}
par(mfrow=c(2,2), mar=c(5,4,1,1))
SSplotSelex(base.model1.1,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.1", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.2,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.2", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.3,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.3", side=3, line=0.5, cex=1.2)
SSplotSelex(base.model1.4,
            subplots = 1,
            mainTitle = FALSE)
mtext("s1.4", side=3, line=0.5, cex=1.2)
```

This Figure \@ref(fig:index2) shows standardized time series of input indices used in four different model scenarios (s1.1 to s1.4) for the stock assessment of Antarctic krill in Subarea 48.1. Each panel presents fishery-dependent (FISHERY) and fishery-independent (SURVEY) indices across five management strata: Bransfield Strait (BS), Elephant Island (EI), Gerlache Strait (GS), Joinville Island (JOIN), and South West (SW), from the mid-1990s to 2020. Scenario s1.4 additionally includes a predator index, highlighting the incorporation of ecosystem variables. The figure illustrates spatial and temporal variability in data availability and trends among sources.

```{r index2, fig.height=8, fig.width=8, fig.cap= "Standardized indices of krill abundance used as input in four model scenarios (s1.1 to s1.4), representing fishery-dependent (FISHERY) and fishery-independent (SURVEY) data across five spatial strata: Bransfield Strait (BS), Elephant Island (EI), Gerlache Strait (GS), Joinville Island (JOIN), and South West (SW). Scenario s1.4 also incorporates a predator index (PREDATOR), reflecting the integration of ecosystem variables into the assessment framework"}
par(mfrow=c(2,2), mar=c(5,4,2,1)) 
SSplotIndices(base.model1.1, subplots = 9)
mtext("s1.1", side=3, line=0.5, cex=1.2)
SSplotIndices(base.model1.2, subplots = 9)
mtext("s1.2", side=3, line=0.5, cex=1.2)
SSplotIndices(base.model1.3, subplots = 9)
mtext("s1.3", side=3, line=0.5, cex=1.2)
SSplotIndices(base.model1.4, subplots = 9)
mtext("s1.4", side=3, line=0.5, cex=1.2)
```

### Outputs Variables `s1.1`

# addWorksheet(wb, "variable_s2")
# writeData(wb, "variable_s2", outps1)
# 
# # Guardar el workbook
# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
# 
# 
# out_s2 <- read_excel("DataKrill.xlsx", 
#     sheet = "variable_s2")
outps1 %>%
  kbl(booktabs = TRUE,
      format = "html",
    caption = "Main variables outputs from stock asssessment krill in WAP `s1.1`") %>%
 kable_styling(latex_options = c("scale_down",
                                    "striped"),# Expande la tabla al ancho completo
                font_size = 8)  #%>% 
  #pack_rows(index = c("Estimation" = 1,
   #                     "Prediction" = 45))
```

### Outputs Variables `s1.2`

# addWorksheet(wb, "variable_s2")
# writeData(wb, "variable_s2", outps1)
# 
# # Guardar el workbook
# saveWorkbook(wb, "DataKrill.xlsx", overwrite = TRUE)
# 
# 
# out_s2 <- read_excel("DataKrill.xlsx", 
#     sheet = "variable_s2")
outps2 %>%
  kbl(booktabs = TRUE,
      format = "html",
    caption = "Main variables outputs from stock asssessment krill in WAP in `s1.2`") %>%
    kable_styling(latex_options = c("scale_down",
                                    "striped"),# Expande la tabla al ancho completo
                font_size = 8) #%>% 
  #pack_rows(index = c("Estimation" = 1,
   #                     "Prediction" = 45))
```

### Outputs Variables `s1.3`
